ANALYZING Nephila male VO2, performance, and  predicted fitness
===============================================================

General housekeeping
--------------------

First load the libraries that we are using
```{r}
library(lme4)
library(lmerTest)
library(ggplot2)
library(knitr)
library(plyr)
library(rptR)
library(PerformanceAnalytics)
library(MuMIn)
```

Next load the functions that we'll use for summary data

Using the script written here: http://www.r-bloggers.com/more-on-exploring-correlations-in-r/
to create the correlation data for the table, but also to create a figure of correlation values

See http://goo.gl/nahmV for documentation of this function
```{r}
cor.prob <- function (X, dfr = nrow(X) - 2) {
  R <- cor(X, use="pairwise.complete.obs")
  above <- row(R) < col(R)
  r2 <- R[above]^2
  Fstat <- r2 * dfr/(1 - r2)
  R[above] <- 1 - pf(Fstat, 1, dfr)
  R[row(R) == col(R)] <- NA
  R
}
```

Use this function to dump the cor.prob output to a 4 column matrix with row/column indices, correlation, and p-value. See StackOverflow question: http://goo.gl/fCUcQ

```{r}
flattenSquareMatrix <- function(m) {
  if( (class(m) != "matrix") | (nrow(m) != ncol(m))) stop("Must be a square matrix.") 
  if(!identical(rownames(m), colnames(m))) stop("Row and column names must be equal.")
  ut <- upper.tri(m)
  data.frame(i = rownames(m)[row(m)[ut]],
             j = rownames(m)[col(m)[ut]],
             cor=t(m)[ut],
             p=m[ut])
}
```

Use this one for summary statistics

```{r}
my.summary <- function(x, na.rm=TRUE){
  result <- c(Mean=mean(x, na.rm=na.rm),
              SD=sd(x, na.rm=na.rm),
              Median=median(x, na.rm=na.rm),
              Min=min(x, na.rm=na.rm),
              Max=max(x, na.rm=na.rm), 
              N=length(x))
}
```


This is for creating the multiplot
 Multiple plot function

 ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
 - cols:   Number of columns in layout
 - layout: A matrix specifying the layout. If present, 'cols' is ignored.

 If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
 then plot 1 will go in the upper left, 2 will go in the upper right, and
 3 will go all the way across the bottom.


```{r}

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

Load the data for the performance analysis (Spider VO2 Data.csv).

```{r}
data = read.table(file.choose(),  na.strings =".", header=T, sep=",")
```

Delete MV37, SV12 & SV14 as they were missing legs after the metabolic analysis and we don't have measurements for MV37 and don't have physiological measurements as we stopped them because they lost enough legs (SV12, SV14)

```{r}
data<-data[-which(is.na(data$Ceph)),]
data<-data[-which(is.na(data$Rest_slope)),]
```

Delete Rows that are missing values in 'Trial' to subset individuals that were used in the mate choice trials

```{r}
data2<-data[-which(is.na(data$trial)),]
```

First scale all the independent variables to create z-scores

```{r}
data2$zleg <- scale(data2$leg, center = TRUE, scale=TRUE)
data2$zceph <- scale(data2$Ceph, center = TRUE, scale=TRUE)
data2$zweight <- scale(data2$Weight, center = TRUE, scale=TRUE)
data2$zvolspeedmax <- scale(data2$Vol_speed_max, center = TRUE, scale=TRUE)
data2$zforcedspeedmax <- scale(data2$Forced_speed_max, center = TRUE, scale=TRUE)
data2$zresting <- scale(data2$Resting, center = TRUE, scale=TRUE)
data2$zactive <- scale(data2$Active, center = TRUE, scale=TRUE)
data2$zscope <- scale(data2$Scope, center = TRUE, scale=TRUE)
data2$zvolspeedmean <- scale(data2$Vol_speed_mean, center = TRUE, scale=TRUE)
data2$zforcedspeedmean <- scale(data2$Forced_speed_mean, center = TRUE, scale=TRUE)
```

Load the data for the repeatability analysis (Spider_Repeatability.csv).

```{r}
perf = read.table(file.choose(),  na.strings =".", header=T, sep=",")
```


First scale all the independent variables to create z-scores

```{r}
perf$zleg <- scale(perf$leg, center = TRUE, scale=TRUE)
perf$zceph <- scale(perf$Ceph, center = TRUE, scale=TRUE)
perf$zweight <- scale(perf$Weight, center = TRUE, scale=TRUE)
perf$zvolspeedmean <- scale(perf$Vol_speed_mean, center = TRUE, scale=TRUE)
perf$zvolspeedmax <- scale(perf$Vol_speed_max, center = TRUE, scale=TRUE)
perf$zforcedspeedmean <- scale(perf$Forced_speed_mean, center = TRUE, scale=TRUE)
perf$zforcedspeedmax <- scale(perf$Forced_speed_max, center = TRUE, scale=TRUE)
```

Data to check differences between replicates (all_males.csv).

```{r}
diff = read.table(file.choose(),  na.strings =".", header=T, sep=",")
```


Exploring the performance data
------------------------------

Check to see if there is a difference in the morphological or physiological variables of individuals that did or did not survive until the next day.

```{r}
DeathCeph <- aov(Ceph ~ Died, data=data)
summary(DeathCeph)

DeathLeg <- aov(leg ~ Died, data=data)
summary(DeathLeg)

DeathWeight <- aov(Weight ~ Died, data=data)
summary(DeathWeight)

DeathScope <- aov(Scope ~ Died + Weight, data=data)
summary(DeathScope)
```

get the columns that you want from the dataset
then create a correlation matrix
then get the p-values using the above function
"flatten"" theat table using the above function

```{r}
corrdata2 <- data2[, c(27,29,34,30,31)]
cor(corrdata2)
cor.prob(corrdata2)
flattenSquareMatrix(cor.prob(corrdata2))
```

Finally, plot the data for Figure S3
```{r}
chart.Correlation(corrdata2)
pdf(file="test.pdf")

```

Using a MANOVA to examine whether any of the performance variables are affected by morphological (size, weight) and physiological (scope) traits.

```{r}
performance_running1 <- manova(cbind(zvolspeedmean, zforcedspeedmax) ~ zscope + zleg + zweight, data=data2)
summary(performance_running1, test="Pillai")

```

No significance of anything.
Below are univariate results just to see if there is any other effect

```{r}
summary.aov(performance_running1)
```

It seems that none of the factors predict any aspect of performance

Now check to see if any of the morphological, physiological, or performance variables affect whether a male found a female or the estimated fitness

But will run a GLM with trial as a random effect as males shared a trial and another males success may have been dependent upon another male

```{r}
Found <- glmer(Found_female ~ zvolspeedmean + zforcedspeedmax + zscope + zleg + zweight + (1|trial), data = data2, family=binomial, control = glmerControl(optimizer = "bobyqa"))

summary(Found)
```

Only max voluntary speed is significant. This does make sense that voluntary speed predicts likelihood of finding a mate.

Now let's check if the amount of time it took to find a female is predicted by anything
```{r}
Time <- lmer(Time_find ~ zvolspeedmean + zforcedspeedmax + zscope + zleg + zweight + (1|trial), data = data2)

summary(Time)
```

Now check to see if estimated paternity (based on the distance to the female and the relative number of rivales) is determined by any of the above variables

```{r}
Fitness <- lmer(Order ~ zvolspeedmean + zforcedspeedmax + zscope + zleg + zweight + (1|trial), data = data2)

summary(Fitness)
```

voluntary running speed is nearly significant.

On average, what was the proportion of their max speed were individuals going? and did that differ between individuals that did and did not find a female?

```{r}
data2$PropMax <- (data2$Vol_speed_mean/data2$Forced_speed_max)

mean(data2$PropMax)
sd(data2$PropMax)/sqrt(length(data2$PropMax))

PropMaxDiff <- lm(PropMax~Found_female, data=data2)
summary(PropMaxDiff)

PropMaxDiff2 <- lm(PropMax~Est_paternity, data=data2)
summary(PropMaxDiff2)


meanProp <- ddply(data2, .(Found_female), summarize, mean=mean(PropMax), se=sd(PropMax)/sqrt(length(PropMax)))
meanProp
```

Is the proportion of their max performance determined by their scope?

```{r}
PropMaxPerf <- lm(PropMax ~ zscope + zweight, data=data2)
summary(PropMaxPerf)
```

Nope

plotting some of the factors here.

First voluntary running speed and searching success
First recode the y/n into 1/0
This makes Figure 1

```{r}
data2$Found2 <- revalue(data2$Found_female, c("Y"="1", "N"="0"))

Figure1 <- ggplot(data2, aes(x=Found_female, y=Vol_speed_max)) + geom_violin() + guides(fill=FALSE) + 
  xlab("Found a female") +
  ylab("Maximum voluntary speed") +
  stat_summary(fun.y="mean", geom="point", size=3) +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())

Figure1

ggsave(
  "Figure1.png",
  Figure1,
  width = 5.25,
  height = 5.25,
  dpi = 1200
)
```

Let's take a look at the relationship between voluntary and forced performance and plot it
This makes Figure 3.

```{r}
MaxPerf <- lm(Vol_speed_mean ~ Forced_speed_max, data=data2)
summary(MaxPerf)

Figure2 <- ggplot(data2, aes(x = Forced_speed_max, y = Vol_speed_mean)) + geom_point(shape=1, size=4) +
  theme_bw() + geom_smooth(method=lm, se=TRUE, fullrange=T) + 
  xlab("Maximum forced running speed (mm "~s^-1*")") +
  ylab("Voluntary running speed (mm "~s^-1*")") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=20,face="bold"),
        axis.title.y=element_text(vjust=0.25),
        axis.title.x=element_text(vjust=0.25),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())

Figure2
        
ggsave(
  "Figure2.png",
  Figure2,
  width = 8.25,
  height = 5.25,
  dpi = 1200
)
```


Exploring the repeatability data
--------------------------------

First, use a MANOVA to examine whether any of the performance variables are affected by size and weight.

```{r}
perf_repeat_vol <- lmer(zvolspeedmean ~ zleg + zweight + (1|ID), data = perf)
summary(perf_repeat_vol)

perf_repeat_forcedmax <- lmer(zforcedspeedmax ~ zleg + zweight + (1|ID), data = perf)
summary(perf_repeat_forcedmax)

```


Leg size and weight predicts voluntary mean running speed

```{r}
Figure2a <- ggplot(perf, aes(x = Weight, y = Vol_speed_mean)) + geom_point(shape=1, size=4) +
  theme_bw() + geom_smooth(method=lm, se=TRUE, fullrange=T) + 
  xlab("Mass (mg)") +
  ylab("Voluntary running speed (mm "~s^-1*")") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=18),
        axis.title.y=element_text(vjust=0.5),
        axis.title.x=element_text(vjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())

Figure2a

Figure2b <- ggplot(perf, aes(x = Vol_speed_mean, y = leg)) + geom_point(shape=1, size=4) +
  theme_bw() + geom_smooth(method=lm, se=TRUE, fullrange=T) + 
  xlab("Size(mm)") +
  ylab("Voluntary running speed (mm "~s^-1*")") +
  theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=18),
        axis.title=element_text(size=18),
        axis.title.y=element_text(vjust=0.5),
        axis.title.x=element_text(vjust=0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())

Figure2b

multiplot(Figure2a, Figure2b, cols=2)

```

What is the proportion of their max speed were individuals going in this second sample of males?

```{r}
perf$PropMax <- (perf$Vol_speed_mean/perf$Forced_speed_max)

mean(perf$PropMax)
sd(perf$PropMax)/sqrt(length(perf$PropMax))

MaxPerf_rpt <- lmer(Vol_speed_mean ~ Forced_speed_max + (1|ID), data=perf)
summary(MaxPerf_rpt)

r.squaredGLMM(MaxPerf_rpt)

```

```{r}
volspeed_rpt <- rpt(zvolspeedmean ~ (1 | ID), grname = "ID", data = perf, datatype = "Gaussian", 
    nboot = 1000, npermut = 0)

summary(volspeed_rpt)
print(volspeed_rpt) 
plot(volspeed_rpt, cex.main = 1)

volmax_rpt <- rpt(zvolspeedmax ~ (1 | ID), grname = "ID", data = perf, datatype = "Gaussian", 
    nboot = 1000, npermut = 0)

summary(volmax_rpt)
print(volmax_rpt) 
plot(volmax_rpt, cex.main = 1)

forced_rpt <- rpt(zforcedspeedmean ~ (1 | ID), grname = "ID", data = perf, datatype = "Gaussian", nboot = 1000, npermut = 0)

summary(forced_rpt)
print(forced_rpt)
plot(forced_rpt, cex.main = 1)

forced_max_rpt <- rpt(zforcedspeedmax ~ (1 | ID), grname = "ID", data = perf, datatype = "Gaussian", nboot = 1000, npermut = 0)

summary(forced_max_rpt)
print(forced_rpt)
plot(forced_max_rpt, cex.main = 1)

```


Checking for differences between males in the two experiments
=============================================================


```{r}
diff1 <- aov(Experiment ~ leg + Weight, data=diff)
summary(diff1)

```
